---
title: "Predicting the Outcome of Armed Conflict"
subtitle: "Raymond Santiago Flores"
format: html
---

### Loading Librarys

```{r, warning=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(astsa)
library(forecast)
library(readxl)
library(conflicted)
library(tidyverse) 
library(cshapes)
library(WDI)
```

### Loading Conflict Data

```{r, warning=FALSE}
# Core conflict data
conflict      <- readRDS("Data/UcdpPrioConflict_v25_1.rds")

# Battle deaths
deaths_conf   <- readRDS("Data/BattleDeaths_v25_1_conf.rds")
deaths_dyad   <- readRDS("Data/BattleDeaths_v25_1.rds")

# GED events
ged           <- readRDS("Data/GEDEvent_v25_1.rds")

# Termination data
termination      <- read_csv("Data/UCDPConflictTerminationDataset_v4_2024_Conflict.csv")
termination_dyad <- read_csv("Data/UCDPConflictTerminationDyad.csv")

# Spreadsheets
issues <- read_excel("Data/Conflict Issues Dataset.xlsx")
view(issues)
esd    <- read_excel("Data/External Support Dataset (ESD).xlsx")
peace  <- read_excel("Data/Peace Agreement.xlsx")
```

### Contry Demographics via WDI

```{r, warning=FALSE}
world_bank <- WDI::WDI(
    indicator = c(
        gdp_per_cap = "NY.GDP.PCAP.KD", #GDP per capita
        population = "SP.POP.TOTL", #Total population
        military_spend = "MS.MIL.XPND.GD.ZS" #Military expenditure % of GDP
    ),
    start = 1960,
    end = 2024,
    extra = FALSE
)

View(world_bank)
```

### Geographic/Terrain data via cshapes

```{r, warning=FALSE}

install.packages("cshapes")
library(cshapes)
shapes <- cshp(date = as.Date("2019-01-01"))




```

### Democracy Scores via vdemdata

```{r, warning=FALSE}
# Still considering the importance of this,

```

### Joining Data 

```{r}
install.packages("countrycode")
library(countrycode)
# Example: add ISO codes to world_bank for easier joining
world_bank <- world_bank |>
  mutate(gwno = countrycode(iso2c, origin = "iso2c", destination = "gwn"))
summary(world_bank)
```

Looking Next

### **Viz 3 — Conflict Issues Heatmap**

```{r}
colnames(conflict)

ggplot()

```

### **Viz 4 — Escalation Patterns**

```{r}

```

### **Viz 5 — External Support Scatter Tool: Plotly via ggplotly()**

```{r}

```

### **Viz 6 - Conflict Termination Alluvial**

```{r}

```

Code for Random Forest model

#### Assumptions check/loading package

```{r}
#Packages
library(randomForest)
#Assumptions check

#Preliminary Code for random forest


#Visualizations for a random forest

```


### **Viz 7 - Intelligence Dashboard**

```{r}

```

### **Viz 8 - Conflict Outcome Predictor**

```{r}

```



